Self-tunable approximated explicit MPC: Heat exchanger implementation and analysis

### 2023-06-22 paper/ATE/main.tex
ATE update4

			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any further tuning or interventions of the control engineer during operation. Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. It provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning process. The designed self-tunable controller improved control performance by decreasing integral squared error, maximal overshoots/undershoots and settling time compared to the conventional control strategy based on a single controller. 
			\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%Research highlights
		\begin{highlights}
			% 3-5 highlights, every one should not exceed 85 characters including spaces
			\item Self-tunable control technique does not need any intervention during control
			\item Tuning based on operating conditions compensates for the nonlinear plant behavior
			\item Implementation of the proposed tuning method on a laboratory heat exchanger 
			\item Control performance improved compared to utilizing explicit MPC with constant setup
			\end{highlights}

@@

	The heat exchangers in their numerous variants are integrated into the many industrial plants as the heat transfer represents the crucial phenomena for all thermal energy applications~\cite{KV18}. Simultaneously, the heat exchangers are energy-demanding and resource-intensive plants. Moreover, from the control viewpoint, the controller design taking into account the nonlinear and asymmetric behavior of the heat exchangers is a challenging task, see~\cite{RL20}. Although the conventional and widely-used proportional–integral–derivative (PID) controllers are robust and easy to tune, their control performance may not be sufficient when it comes to energy savings. Various extensions built above the well-tuned PID controller were developed to compensate for the nonlinear and asymmetric behavior, often affected by the additional impact of the uncertain parameters originated, e.g., in the fouling~\cite{MT19}. Such widely-used
	control strategies include, e.g., the robust control~\cite{WY18}, the gain-scheduling, adaptive and optimal control. For the rigorous mathematical modeling and controller design methods in general, see~\cite{MF08} and for the controller design tailored for the process control engineers see~\cite{Liptak}.

	One of the promising control strategies addressing all these issues in an optimal way came with the formation of the model predictive control (MPC), e.g., see~\cite{Morari_MPC}. MPC provides optimal control input based on the minimization of a specified cost function while considering a model of the system. Compared to linear quadratic controllers (LQR)~\cite{LQR}, model predictive control also includes constraints on the control input or process variables~\cite{Maciejowski_MPC}, and additional saturation is not necessary. Moreover, as the optimization problem is solved in every control step, MPC represents a receding horizon control policy~\cite{receding_horizon}, having a significant benefit mainly in the terms of disturbance rejection. The model predictive control was intensively investigated in connection with heat exchangers. In~\cite{Vinaya_HE_MPC}, the authors developed a model predictive control for a shell and tube heat exchanger. In~\cite{Oravec_energy}, the control performance of the robust MPC implemented for a plate heat exchanger with soft constraints was analyzed. Four robust control strategies were presented and compared in~\cite{Oravec_HE_ATE}. A two-level control structure was applied on a heat exchanger network in~\cite{Gonzales_HE_MPC}, where the low level of control was ensured by MPC and the high level by a supervisory online optimizer. The fast nonlinear MPC was designed to optimize the waste heat recovery~\cite{WC19}.  The multi-layer control designed in~\cite{DZ18} designed the MPC in the leader loop to optimize the thermal response to improved control performance.

@@

	Explicit model predictive control~\cite{Bemporad_automatica} utilizes a parametric solution of the model predictive control introducing its application range towards the systems with fast dynamics. Moreover, the explicit solution enables providing rigorous analysis and certification of the closed-loop system stability, constraints satisfaction, etc. As the explicit solution is available, real-time solving of the optimization problem in every control step is omitted. As this work deals with industrial-oriented implementation, let us consider the optimization problem in the following form:
	\begin{subequations}
		\label{eq:mpc_problem}
		\begin{eqnarray}
			\label{eq:mpc_problem_cost}
			\min_{u_0,u_{1},\ldots,u_{N-1}} &~& \! \sum_{k=0}^{N-1} \! \left( (y_k-y_\mathrm{ref})^{\intercal} Q_\mathrm{y} (y_k-y_\mathrm{ref}) + u_{k}^{\intercal} R u_{k} + x_{\mathrm{I},k}^{\intercal} Q_\mathrm{x} x_{\mathrm{I},k} \right)  \\
			\label{eq:mpc_problem_prediction_model_x}
			\mathrm{s.t.\!:} &~& \widetilde{x}_{k+1} = \widetilde{A}\,\widetilde{x}_{k} + \widetilde{B}\,u_{k}, \\
			\label{eq:mpc_problem_prediction_model_y}
			&~& y_{k} = \widetilde{C}\,\widetilde{x}_{k}, \\
			\label{eq:mpc_problem_input_constraints}
			&~& u_{k} \in \mathcal{U}, \\
			\label{eq:mpc_problem_state_constraints}
			&~& y_{k} \in \mathcal{Y}, \\
			\label{eq:mpc_problem_initial_coindition}
			&~& \widetilde{x}_{0} = \theta, \\
			\label{eq:mpc_problem_k_range}
			&~& k = 0,1,\ldots, N-1,
		\end{eqnarray}
	\end{subequations}
	where $k$ denotes the step of the prediction horizon $N$. 
	%
	To obtain the offset-free control results, the built-in integrator was included in the state-space model, e.g., see~\cite{Ruscio_MPC_integral}. 
	The prediction model in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} has the form of augmented linear time-invariant (LTI) system for a given augmented state matrix $\widetilde{A} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\widetilde{\mathrm{x}}}}$, augmented input matrix $\widetilde{B} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\mathrm{u}}}$ and augmented output matrix $\widetilde{C} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\widetilde{\mathrm{x}}}}$. 
	%
	Variables $\widetilde{x} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}}}$, $u \in \mathbb{R}^{n_{\mathrm{u}}}$, $y \in \mathbb{R}^{n_{\mathrm{y}}}$ are vectors of corresponding augmented system states, control inputs, and system outputs, respectively. 
	%
	The sets $\mathcal{U} \subseteq \mathbb{R}^{n_{\mathrm{u}}}$, $\mathcal{Y} \subseteq \mathbb{R}^{n_{\mathrm{y}}}$ are convex polytopic sets of physical constraints on inputs and outputs, respectively. The penalty matrix $Q_\mathrm{y} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}$ penalizes the squared control error, i.e., the deviation between the output and output reference value $y_\mathrm{ref}$. The matrix $R \in \mathbb{R}^{n_{\mathrm{u}} \times n_{\mathrm{u}}}$ penalizes the squarred value of control inputs. 
	%
	The value of integrator is also penalized in the cost function with the penalty matrix $Q_\mathrm{x} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}$. 
	%
	The parameter $\theta \in \Theta$ in Eq.~\eqref{eq:mpc_problem_initial_coindition} represents the initial condition of the optimization problem for which it is parametrically pre-computed.

	The augmented model of the controlled system with the built-in integrator in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} us rewritten as follows:
	\begin{subequations}
		\begin{eqnarray} 
			\label{eq:mpc_augmented_model_x} 
			\widetilde{x}_{k+1} &=& \begin{bmatrix} x_{k+1} \\ x_{\mathrm{I},k+1}\end{bmatrix} = \begin{bmatrix} A & \textit{0} \\ -T_\mathrm{s} C & I \end{bmatrix} \begin{bmatrix} x_{k} \\ x_{\mathrm{I},k} \end{bmatrix} + \begin{bmatrix} B \\ I \end{bmatrix} u_{k}, \\
			\label{eq:mpc_augmented_model_y}
			y_k &=& \begin{bmatrix} C & \textit{0} \end{bmatrix} \begin{bmatrix} x_{k} \\ x_{\mathrm{I},k} \end{bmatrix},
		\end{eqnarray}
	\end{subequations}
	where $x_{\mathrm{I}} \in \mathbb{R}^{n_{\mathrm{y}}}$ is the integral of the control error, $T_\mathrm{s}$ denotes the sampling time, and matrices $A$, $B$, $C$ are well-known the state-space matrices that form the augmented LTI model. As the consequence of this extension and penalization in the cost function in Eq.~\eqref{eq:mpc_problem_cost}, not only the control error is penalized, but also an integrated value, which leads to analogous offset-free reference tracking results as incorporating an integral part in PID controller.

	The parametric solution of the optimization problem of the quadratic programming (QP) in Eq.~\eqref{eq:mpc_problem}, leads to the explicit solution in the form of piecewise affine PWA control law defined above the domain consisting of $r$ critical regions:
	\begin{eqnarray}
		\label{eq:PWA_control_law}
		u(\theta) = \left\{ 
		\begin{matrix}
			F_{1} \, \theta + g_{1} & \mathrm{if} & \quad \theta \in \mathcal{R}_1, \\
			F_{2} \, \theta + g_{2} & \mathrm{else}\,\,\mathrm{if} &\quad \theta \in \mathcal{R}_2, \\
			\vdots & \\
			F_{r} \, \theta + g_{r} & \mathrm{else}\,\,\mathrm{if} & \quad \theta \in \mathcal{R}_{r}, \\
		\end{matrix}
		\right.
	\end{eqnarray}
	where $F_{i} \in \mathbb{R}^{n_{\mathrm{u}} \times n_{\mathrm{x}}}$ and $g_{i}  \in \mathbb{R}^{n_{\mathrm{u}}}$ respectively are the slope and affine section of the corresponding control law. The PWA function defined in Eq.~\eqref{eq:PWA_control_law} is stored and recalled in the online phase, i.e., during the real-time control. Based on identifying the specific critical region $\mathcal{R}_{i}$, where the parameter $\theta$ belongs, the optimal control input is calculated based on the associated control law in Eq.~\eqref{eq:PWA_control_law}.

	Note, many other formulations of the optimization problems for the explicit MPC design, were mainly in the terms of the definition of the cost functions in Eq.~\eqref{eq:mpc_problem_cost}. Also, the incremental (velocity) formulation of the state-space model is common, but leads to further extension of the vector of parameters $\theta$, and therefore also the complexity of the explicit MPC controller increases. Another option for offset-free tracking is introducing disturbance modeling and estimation. For such an overview see e.g.~\cite{Klauco_mpc}

@@

	The aggressivity of the controller and the whole nature of the control is influenced by appropriate fine-tuning of the penalty matrices in the optimization problem in Eq.~\eqref{eq:mpc_problem}. When the multi-parametric QP (mp-QP) problem is precomputed offline to obtain the corresponding parametric solution, it is not possible to tune the controller afterward without trading off a significant increase in the controller complexity or the performance loss. As the operating conditions and requirements of controller setup may differ throughout the control, the ability to adjust the controller's aggressivity can be very beneficial.

@@

	Let us consider the penalty matrices in the cost function in Eq.~\eqref{eq:mpc_problem_cost}. The penalty matrices are scaled in the following way:
	\begin{subequations}
		\label{eq:tunable_matrices}
		\begin{eqnarray}
			\label{eq:tunable_R}
			&~& R(k) = (1-\rho(k)) \, R_\mathrm{L} + \rho(k) \, R_\mathrm{U}, \\
			\label{eq:tunable_Qx}
			&~& Q_\mathrm{x}(k) = (1-\rho(k)) \, Q_\mathrm{x,L} + \rho(k) \, Q_\mathrm{x,U}, \\
			\label{eq:tunable_Qy}
			&~& Q_\mathrm{y}(k) = (1-\rho(k)) \, Q_\mathrm{y,L} + \rho(k) \, Q_\mathrm{y,U},
		\end{eqnarray}
	\end{subequations}
	where $\rho$ represents the tuning parameter such that $0 \le \rho \le 1$ holds. Based on the rules in Eq.~\eqref{eq:tunable_matrices}, it is possible to choose online any controller setup from the lower to the upper boundary of the tuned matrix. From the implementation point of view, it is preferred to tune just a single penalty matrix, i.e., to store only two controllers corresponding to the boundary values of the selected penalty matrix. To determine which penalty matrix in Eq.~\eqref{eq:tunable_matrices} should be tuned, it is suggested to judge the control performance subject to the specific system and systematically tune all of the penalty matrices.

@@

	Based on Eq.~\eqref{eq:rho_self_tunable}, the property $0 \le \rho \le 1$ holds, as $\vert y_{\mathrm{ref}} \vert \le d_{\max}$. As a consequence, the parameter $\rho$ represents a way how to normalize the deviation from the steady-state value and is exploited to scale the penalty matrix or, implicitly, to scale the aggressiveness of the control action.

@@

	\begin{definition}[Scaling of the tuning parameter]
		\label{def:rho_tilde}
		Given the value of tuning parameter $\rho$, $0 \leq \rho \leq 1$, the splitting value of the tuning parameter interval $\rho_{\mathrm{s}}$, $0 < \rho_{\mathrm{s}} < 1$, and the value of the decision function $\gamma$, $0 < \gamma < 1$. Then the scaling of the tuning parameter $\widetilde{\rho}$ is given by:		
		\begin{eqnarray}
			\label{eq:rho_tilde_def}
			\widetilde{\rho} = \left\{ 
			\begin{matrix}
				\rho \, \rho_{\mathrm{s}} & \mathrm{if} & \quad \gamma \in \langle 0, \rho_{\mathrm{s}} \rangle , \\
				\rho \, (1-\rho_{\mathrm{s}}) + \rho_{\mathrm{s}}, & \mathrm{else}\,\,\mathrm{if} &\quad \gamma \in \langle \rho_{\mathrm{s}}, 1 \rangle.
			\end{matrix}
			\right.
		\end{eqnarray}
	\end{definition}

@@

	The following outcomes result from Eq.~\eqref{eq:rho_tilde_def}. 
	%	\begin{corollary}
		% The following outcomes result from Eq.~\eqref{eq:rho_up}. If $\rho=0$, then $\widetilde{\rho} = 0$. When considering the other extreme, i.e., $\rho=1$, then $\widetilde{\rho} = \rho_{\mathrm{s}}$. In such a way, the parameter $\rho$ can be scaled in the first part of the interval up to the splitting value. Note, as the relation in Eq.~\eqref{eq:rho_up} is linear and $0 \le \rho \le 1$ holds, for $\widetilde{\rho}$ it is impossible to acquire other values than in the given interval of parameter $\rho$.		
		%	\end{corollary}
	%
	%	\begin{corollary}
		%		The following outcomes result from Eq.~\eqref{eq:rho_down}. When $\rho=0$, $\widetilde{\rho} = \rho_{\mathrm{s}}$. When considering the other extreme, i.e., $\rho=1$, then $\widetilde{\rho} = 1$. In such a way, the parameter $\rho$ can be scaled in the second part of the interval from the splitting value up to 1. Note, as the relation in Eq.~\eqref{eq:rho_down} is linear and $0 \le \rho \le 1$ holds, for $\widetilde{\rho}$ it is impossible to acquire other values than in the given interval of parameter $\rho$.
		%	\end{corollary}
	\begin{lemma}
		\label{lem:PWA_control_law_interval}
		Given control law in~\eqref{eq:PWA_control_law}, its approximation given by the convex combination in~\eqref{eq:tunable_u}, and given scaled tuning parameter $\widetilde{\rho}$ according Definition~\ref{def:rho_tilde}. Then the control law  approximated into the form:
		\begin{eqnarray}
			\label{eq:PWA_control_law_interval}
			u(k) = (1-\widetilde{\rho}(k)) \, u_\mathrm{L}(k) + \widetilde{\rho}(k) \, u_\mathrm{U}(k),
		\end{eqnarray}
		preserves the closed-loop system stability and recursive feasibility of the original control law in~\eqref{eq:PWA_control_law}.
	\end{lemma}

>>>>

			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any interventions of the control engineer during operation \textcolor{red}{in order to properly tune the controller}. Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. It provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning process. The designed self-tunable controller improved control performance by decreasing integral squared error, maximal overshoots/undershoots and settling time compared to the conventional control strategy based on a single controller. 
			\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%Research highlights
		\begin{highlights}
			% 3-5 highlights, every one should not exceed 85 characters including spaces
			\item Self-tunable control technique does not need any intervention during control
			\item Tuning based on operating conditions compensates for the nonlinear plant behavior
			\item Implementation of the proposed tuning method on a laboratory heat exchanger 
			\item Control performance improved compared to utilizing explicit MPC with constant setup
			\end{highlights}

@@

	The heat exchangers in their numerous variants are integrated into many industrial plants as the heat transfer represents the crucial phenomena for all thermal energy applications~\cite{KV18}. Simultaneously, the heat exchangers are energy-demanding and resource-intensive plants. Moreover, from the control viewpoint, the controller design taking into account the nonlinear and asymmetric behavior of the heat exchangers is a challenging task, see~\cite{RL20}. What should be also considered, is a choice of the controlled variable, as the outlet stream temperature does not have to be always the efficient one. In~\cite{JIN_CV}, the idea is to optimize the controlled variable of counter flow heat exchanger. A new controlled variable named as heat exchanger comprehensive effectiveness is introduced, based on maximizing the amount of heat exchange and minimizing the irreversible loss.

	Very common challenge in terms of heat exchangers is fouling. In~\cite{TRAFCZYNSKI_fouling}, the authors address this effect while using proportional–integral–derivative (PID) control. If the thermal resistance of fouling is increased and the parameters of controller remain unchanged, it could lead to oscillations of the controlled temperature, too slow response to setpoint changes or significant temperature overshoot. To prevent these effects, the PID controller parameters are adjusted.

	Although the conventional and widely-used PID controllers are robust and easy to tune, their control performance may not be sufficient when it comes to energy savings. Various extensions built above the well-tuned PID controller were developed to compensate for the nonlinear and asymmetric behavior, often affected by the additional impact of the uncertain parameters originated, e.g., in the already mentioned fouling~\cite{MT19}. Such widely-used control strategies include, e.g., the robust control~\cite{WY18}, the gain-scheduling, adaptive and optimal control. In~\cite{VASICKANINOVA_HE}, a neural network predictive controller together with the fuzzy controller in the complex control structure with an auxiliary control variable was investigated. For the rigorous mathematical modeling and controller design methods in general, see~\cite{MF08} and for the controller design tailored for the process control engineers see~\cite{Liptak}.

	One of the promising control strategies addressing all these issues in an optimal way came with the formation of the model predictive control (MPC), e.g., see~\cite{Morari_MPC}. MPC provides optimal control input based on the minimization of a specified cost function while considering a model of the system. Compared to linear quadratic controllers (LQR)~\cite{LQR}, model predictive control also includes constraints on the control input or process variables~\cite{Maciejowski_MPC}, and additional saturation is not necessary. Moreover, as the optimization problem is solved in every control step, MPC represents a receding horizon control policy~\cite{receding_horizon}, having a significant benefit mainly in the terms of disturbance rejection. The model predictive control was intensively investigated in connection with heat exchangers. In~\cite{Vinaya_HE_MPC}, the authors developed a model predictive control for a shell and tube heat exchanger. %In~\cite{Oravec_energy}, the control performance of the robust MPC implemented on a plate heat exchanger with soft constraints was analyzed. 
	Four robust control strategies were presented and compared in~\cite{Oravec_HE_ATE}. A two-level control structure was applied on a heat exchanger network in~\cite{Gonzales_HE_MPC}, where the low level of control was ensured by MPC and the high level by a supervisory online optimizer. The fast nonlinear MPC was designed to optimize the waste heat recovery~\cite{WC19}. The multi-layer control designed in~\cite{DZ18} designed the MPC in the leader loop to optimize the thermal response to improved control performance. In~\cite{DYRSKA_HE}, the authors focus on reduction of computational complexity associated with the evaluation of control action of MPC. By implementing the constraint removal technique, the energy consumption of the microcontroller is reduced.

@@

	Explicit model predictive control~\cite{Bemporad_automatica} utilizes a parametric solution of the model predictive control introducing its application range towards the systems with fast dynamics. Moreover, the explicit solution enables providing rigorous analysis and certification of the closed-loop system stability, constraints satisfaction, etc. As the explicit solution is available, real-time solving of the optimization problem in every control step is omitted. As this work deals with industrial-oriented implementation, let us consider the optimization problem in the following form:
	\begin{subequations}
		\label{eq:mpc_problem}
		\begin{eqnarray}
			\label{eq:mpc_problem_cost}
			\min_{u_0,u_{1},\ldots,u_{N-1}} &~& \! \sum_{k=0}^{N-1} \! \left( (y_k-y_\mathrm{ref})^{\intercal} Q_\mathrm{y} (y_k-y_\mathrm{ref}) + u_{k}^{\intercal} R u_{k} + x_{\mathrm{I},k}^{\intercal} Q_\mathrm{I} x_{\mathrm{I},k} \right)  \\
			\label{eq:mpc_problem_prediction_model_x}
			\mathrm{s.t.\!:} &~& \widetilde{x}_{k+1} = \widetilde{A}\,\widetilde{x}_{k} + \widetilde{B}\,u_{k}, \\
			\label{eq:mpc_problem_prediction_model_y}
			&~& y_{k} = \widetilde{C}\,\widetilde{x}_{k}, \\
			\label{eq:mpc_problem_input_constraints}
			&~& u_{k} \in \mathcal{U}, \\
			\label{eq:mpc_problem_state_constraints}
			&~& y_{k} \in \mathcal{Y}, \\
			\label{eq:mpc_problem_initial_coindition}
			&~& \widetilde{x}_{0} = \theta, \\
			\label{eq:mpc_problem_k_range}
			&~& k = 0,1,\ldots, N-1,
		\end{eqnarray}
	\end{subequations}
	where $k$ denotes the step of the prediction horizon $N$. 
	%
	To obtain the offset-free control results, the built-in integrator was included in the state-space model, e.g., see~\cite{Ruscio_MPC_integral}. 
	The prediction model in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} has the form of augmented linear time-invariant (LTI) system for a given augmented state matrix $\widetilde{A} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\widetilde{\mathrm{x}}}}$, augmented input matrix $\widetilde{B} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\mathrm{u}}}$ and augmented output matrix $\widetilde{C} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\widetilde{\mathrm{x}}}}$. 
	%
	Variables $\widetilde{x} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}}}$, $u \in \mathbb{R}^{n_{\mathrm{u}}}$, $y \in \mathbb{R}^{n_{\mathrm{y}}}$ are vectors of corresponding augmented system states, control inputs, and system outputs, respectively. 
	%
	The sets $\mathcal{U} \subseteq \mathbb{R}^{n_{\mathrm{u}}}$, $\mathcal{Y} \subseteq \mathbb{R}^{n_{\mathrm{y}}}$ are convex polytopic sets of physical constraints on inputs and outputs, respectively. The penalty matrix $Q_\mathrm{y} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}$ penalizes the squared control error, i.e., the deviation between the output and output reference value $y_\mathrm{ref}$. The matrix $R \in \mathbb{R}^{n_{\mathrm{u}} \times n_{\mathrm{u}}}$ penalizes the squarred value of control inputs. 
	%
	The value of integrator is also penalized in the cost function with the penalty matrix $Q_\mathrm{I} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}$. 
	%
	The parameter $\theta \in \Theta$ in Eq.~\eqref{eq:mpc_problem_initial_coindition} represents the initial condition of the optimization problem for which it is parametrically pre-computed.

	The augmented model of the controlled system with the built-in integrator in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} is rewritten as follows:
	\begin{subequations}
		\begin{eqnarray} 
			\label{eq:mpc_augmented_model_x} 
			\widetilde{x}_{k+1} &=& \begin{bmatrix} x_{k+1} \\ x_{\mathrm{I},k+1}\end{bmatrix} = \begin{bmatrix} A & \textit{0} \\ -T_\mathrm{s} C & I \end{bmatrix} \begin{bmatrix} x_{k} \\ x_{\mathrm{I},k} \end{bmatrix} + \begin{bmatrix} B \\ I \end{bmatrix} u_{k}, \\
			\label{eq:mpc_augmented_model_y}
			y_k &=& \begin{bmatrix} C & \textit{0} \end{bmatrix} \begin{bmatrix} x_{k} \\ x_{\mathrm{I},k} \end{bmatrix},
		\end{eqnarray}
	\end{subequations}
	where $x_{\mathrm{I}} \in \mathbb{R}^{n_{\mathrm{y}}}$ is the integral of the control error, $T_\mathrm{s}$ denotes the sampling time, and matrices $A$, $B$, $C$ are the well-known state-space matrices that form the augmented LTI model. As a consequence of this extension and penalization in the cost function in Eq.~\eqref{eq:mpc_problem_cost}, not only the control error is penalized, but also the integrated value, which leads to analogous offset-free reference tracking results as incorporating an integral part in the PID controller.

	The parametric solution of the optimization problem of the quadratic programming (QP) in Eq.~\eqref{eq:mpc_problem} leads to the explicit solution in the form of piecewise affine PWA control law defined above the domain consisting of $r$ critical regions:
	\begin{eqnarray}
		\label{eq:PWA_control_law}
		u(\theta) = \left\{ 
		\begin{matrix}
			F_{1} \, \theta + g_{1} & \mathrm{if} & \quad \theta \in \mathcal{R}_1, \\
			F_{2} \, \theta + g_{2} & \mathrm{else}\,\,\mathrm{if} &\quad \theta \in \mathcal{R}_2, \\
			\vdots & \\
			F_{r} \, \theta + g_{r} & \mathrm{else}\,\,\mathrm{if} & \quad \theta \in \mathcal{R}_{r}, \\
		\end{matrix}
		\right.
	\end{eqnarray}
	where $F_{i} \in \mathbb{R}^{n_{\mathrm{u}} \times n_{\mathrm{x}}}$ and $g_{i}  \in \mathbb{R}^{n_{\mathrm{u}}}$ respectively are the slope and affine section of the corresponding control law. The PWA function defined in Eq.~\eqref{eq:PWA_control_law} is stored and recalled in the online phase, i.e., during the real-time control. Based on identifying the specific polytopic critical region $\mathcal{R}_{i}$, where the parameter $\theta$ belongs, the optimal control input is calculated based on the associated control law in Eq.~\eqref{eq:PWA_control_law}.

	Note, many other formulations of the optimization problems for the explicit MPC design were formulated mainly in terms of the definition of the cost functions in Eq.~\eqref{eq:mpc_problem_cost}. Also, the incremental (velocity) formulation of the state-space model is common, but leads to further extension of the vector of parameters $\theta$, and therefore also the complexity of the explicit MPC controller increases. Another option for offset-free tracking is introducing disturbance modeling and estimation. For such an overview see e.g.~\cite{Klauco_mpc}

@@

	The aggressivity of the controller and the whole nature of the control is influenced by appropriate fine-tuning of the penalty matrices in the optimization problem in Eq.~\eqref{eq:mpc_problem}. When the multi-parametric QP (mp-QP) problem is precomputed offline to obtain the corresponding parametric solution, it is not possible to tune the controller afterward without trading off a significant increase in the controller complexity or the performance loss. As the operating conditions and requirements on controller setup may differ throughout the control, the ability to adjust the controller's aggressivity can be very beneficial.

@@

	Let us consider the penalty matrices in the cost function in Eq.~\eqref{eq:mpc_problem_cost}. The penalty matrices are scaled in the following way:
	\begin{subequations}
		\label{eq:tunable_matrices}
		\begin{eqnarray}
			\label{eq:tunable_R}
			&~& R(k) = (1-\rho(k)) \, R_\mathrm{L} + \rho(k) \, R_\mathrm{U}, \\
			\label{eq:tunable_Qx}
			&~& Q_\mathrm{I}(k) = (1-\rho(k)) \, Q_\mathrm{I,L} + \rho(k) \, Q_\mathrm{I,U}, \\
			\label{eq:tunable_Qy}
			&~& Q_\mathrm{y}(k) = (1-\rho(k)) \, Q_\mathrm{y,L} + \rho(k) \, Q_\mathrm{y,U},
		\end{eqnarray}
	\end{subequations}
	where $\rho$ represents the tuning parameter such that $0 \le \rho \le 1$ holds. Based on the rules in Eq.~\eqref{eq:tunable_matrices}, it is possible to choose online any controller setup from the lower to the upper boundary of the tuned matrix. From the implementation point of view, it is preferred to tune just a single penalty matrix, i.e., to store only two controllers corresponding to the boundary values of the selected penalty matrix. To determine which penalty matrix in Eq.~\eqref{eq:tunable_matrices} should be tuned, it is suggested to judge the control performance subject to the specific system and systematically tune all of the penalty matrices.

@@

	Based on Eq.~\eqref{eq:rho_self_tunable}, the property $0 \le \rho \le 1$ holds, as $\vert y_{\mathrm{ref}} \vert \le d_{\max}$. As a consequence, the parameter $\rho$ represents a way how to normalize the deviation from the steady-state value and is exploited to scale the control action or, implicitly, to tune the aggressiveness of the controller.

@@

	\begin{definition}[Scaling of the tuning parameter]
		\label{def:rho_tilde}
		Given the value of tuning parameter $\rho$, $0 \leq \rho \leq 1$, the splitting value of the tuning parameter interval $\rho_{\mathrm{s}}$, $0 < \rho_{\mathrm{s}} < 1$, and the value of the decision function $\gamma$, \textcolor{red}{$0 < \gamma < 1$}. Then the scaling of the tuning parameter $\widetilde{\rho}$ is given by:		
		\begin{eqnarray}
			\label{eq:rho_tilde_def}
			\widetilde{\rho} = \left\{ 
			\begin{matrix}
				\rho \, \rho_{\mathrm{s}} & \mathrm{if} & \quad \gamma \in \langle 0, \rho_{\mathrm{s}} \rangle , \\
				\rho \, (1-\rho_{\mathrm{s}}) + \rho_{\mathrm{s}}, & \mathrm{else}\,\,\mathrm{if} &\quad \gamma \in \langle \rho_{\mathrm{s}}, 1 \rangle.
			\end{matrix}
			\right.
		\end{eqnarray}
	\end{definition}

@@

	The following outcomes result from Eq.~\eqref{eq:rho_tilde_def}. 
	%	\begin{corollary}
		% The following outcomes result from Eq.~\eqref{eq:rho_up}. If $\rho=0$, then $\widetilde{\rho} = 0$. When considering the other extreme, i.e., $\rho=1$, then $\widetilde{\rho} = \rho_{\mathrm{s}}$. In such a way, the parameter $\rho$ can be scaled in the first part of the interval up to the splitting value. Note, as the relation in Eq.~\eqref{eq:rho_up} is linear and $0 \le \rho \le 1$ holds, for $\widetilde{\rho}$ it is impossible to acquire other values than in the given interval of parameter $\rho$.		
		%	\end{corollary}
	%
	%	\begin{corollary}
		%		The following outcomes result from Eq.~\eqref{eq:rho_down}. When $\rho=0$, $\widetilde{\rho} = \rho_{\mathrm{s}}$. When considering the other extreme, i.e., $\rho=1$, then $\widetilde{\rho} = 1$. In such a way, the parameter $\rho$ can be scaled in the second part of the interval from the splitting value up to 1. Note, as the relation in Eq.~\eqref{eq:rho_down} is linear and $0 \le \rho \le 1$ holds, for $\widetilde{\rho}$ it is impossible to acquire other values than in the given interval of parameter $\rho$.
		%	\end{corollary}
	\begin{lemma}
		\label{lem:PWA_control_law_interval}
		Given control law in~\eqref{eq:PWA_control_law}, its approximation given by the convex combination in~\eqref{eq:tunable_u}, and given scaled tuning parameter $\widetilde{\rho}$ according to Definition~\ref{def:rho_tilde}. Then the control action approximated into the form:
		\begin{eqnarray}
			\label{eq:PWA_control_law_interval}
			u(k) = (1-\widetilde{\rho}(k)) \, u_\mathrm{L}(k) + \widetilde{\rho}(k) \, u_\mathrm{U}(k),
		\end{eqnarray}
		preserves the closed-loop system stability and recursive feasibility of the original control law in~\eqref{eq:PWA_control_law}.
	\end{lemma}
